---
title: "README"
output:
  md_document:
    variant: gfm
vignette: >
  %\VignetteIndexEntry{README}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
old <- options(width = 150)
```

Easy-to-use, dependencyless Logger for R
==================================
<!-- badges: start -->
[![CRAN\_Status\_Badge](http://www.r-pkg.org/badges/version/loggit2)](https://cran.r-project.org/package=loggit2)
[![Downloads](https://cranlogs.r-pkg.org/badges/grand-total/loggit2)](https://cran.r-project.org/package=loggit2)
[![R-CMD-check](https://github.com/MEO265/loggit2/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/MEO265/loggit2/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/MEO265/loggit2/graph/badge.svg)](https://codecov.io/gh/MEO265/loggit2)
------------------------------------------------------------------------
<!-- badges: end -->

`loggit2` is an easy-to-use [`ndJSON`](https://github.com/ndjson/ndjson-spec) logging library for R software,
with _zero_ external dependencies.

Please see below for some quick examples, and read the [vignettes](https://cran.r-project.org/web/packages/loggit2/vignettes/) for the
Getting Started guide.

Why use `loggit2`?
-----------------

`loggit2` takes a modern approach to logging in R:

- Opting to use the JSON format
- Highly flexible log streams
- Enables log data analysis on the same host
- _Zero_ external dependencies

Additionally, the boilerplate to get going with `loggit2` is minimal at worst.
No need to write custom formatters, handlers, levels, etc. -- ***just loggit!***

Usage
--------------

The quickest way to get up & running with `loggit2` is to do nearly nothing.

`loggit2` provides a set of wrappings for base R's `message()`, `warning()`, and
`stop()` functions that maintain identical functionality, making it sufficient to
import the `loggit2` namespace, for example by using `library("loggit2")`, or by
prefixing `loggit2::` at the desired locations.

```{r, error = TRUE}
loggit2::message("This is a message")
loggit2::warning("This is a warning")
loggit2::stop("This is an error")
```

You can suppress the additional console output by using `echo = FALSE` and
you won't notice any difference to the base functions (except that the log
will be filled in the background).

```{r, error = TRUE}
base::message("This is another message")
loggit2::message("This is another message", echo = FALSE)

base::warning("This is another warning")
loggit2::warning("This is another warning", echo = FALSE)

base::stop("This is another error")
loggit2::stop("This is another error", echo = FALSE)
```

You can also use `loggit()` directly to compose much more custom
logs, including ***entirely custom fields*** (and prevent throwing actual conditions).
`loggit2` doesn't require that you set custom logger objects or anything like that:
just throw whatever you want at it, and it'll become a structured log.

```{r loggit2}
loggit2::loggit("ERROR", "This will log an error", anything_else = "you want to include")

# Read log file into data frame to implement logic based on entries
loggit2::read_logs()
```

Check out the [vignettes](https://cran.r-project.org/web/packages/loggit2/vignettes/) for more details.

Installation
------------

You can install the latest CRAN release of `loggit2` via

    install.packages("loggit2")

or, get the latest development version from GitHub via

    devtools::install_github("MEO265/loggit2")

Acknowledgments
-----------
This package is based on the ["loggit" package by Ryan Price](https://github.com/ryapric/loggit), specifically version 2.1.1.

Due to technical reasons, this repository is not a GitHub fork of [Ryan's repository](https://github.com/ryapric/loggit).

```{r, include = FALSE}
options(old)
```