---
title: "Further Configurations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Further Configurations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set( #nolint
  collapse = TRUE,
  comment = "#>"
)
old <- options(width = 200L)
```


`loggit2` can be used immediately and fully without configuration after installation. However, to fully utilize the potential of `loggit2`, some configurations can be made.

## Configuration System

`loggit2` uses a three-tier configuration system:

1. Meta/System-wide Configuration
2. Global/Session-wide[^1] Configuration
3. Local/Function-specific Configuration

The first tier is the Meta or System-wide Configuration. It is defined by environment variables that are read by `loggit2` when the package is loaded. These settings then serve as default values for the session-wide configuration.

The second tier is the Global or Session-wide Configuration. It is defined by the `set_*()` functions of `loggit2`.

The third tier is the Local or Function-specific Configuration, implemented through the arguments of the functions. For a quick overview of the functions, refer to the [Getting Started Vignette](loggit2.html).

Each time a function of `loggit2` is called, the settings from the three tiers are combined. A higher tier/specificity configuration always takes precedence over a lower tier/specificity configuration.

## Configuration Options

The configuration options of `loggit2` are described below. For details, refer to the function documentation.

### Log file
The log file can be set either via the environment variable `FILE_LOGGIT2`, the function `set_log_file()`, or the `logfile` argument of the logging functions `loggit()` and `with_loggit()`. The wrappers of the base R condition handler always use the session-wide configuration.


```{r, echo = -1}
setwd(tempdir())
old_log <- loggit2::set_logfile(logfile = "logfile.log")
loggit2::loggit(
  log_lvl = "DEBUG",
  log_msg = "This message will be logged to `logfile.log`."
)


loggit2::loggit(
  log_lvl = "DEBUG",
  log_msg = "This message will be logged to `otherlogfile.log`.",
  logfile = "otherlogfile.log"
)


loggit2::with_loggit(logfile = "logfile2.log", {
  base::message("This message will be logged to `logfile2.log`.")
})

loggit2::set_logfile(old_log)
```

### Log level
The log level can be set either via the environment variable `LEVEL_LOGGIT2` or the function `set_log_level()`. The `.loggit` argument can also be passed to the wrappers of the base R condition handler to enforce logging (`TRUE`) or suppress it (`FALSE`). The logging function `loggit()` can enforce logging through the `ignore_log_level` argument. `with_loggit()` allows setting a log level for the code block being executed.
```{r, error = TRUE}
old_log_lvl <- loggit2::set_log_level("INFO")
loggit2::message("This message will be logged, since the log level is INFO.")
loggit2::loggit(
  log_lvl = "DEBUG",
  log_msg = "This message will not be logged, since the log level is INFO."
)
loggit2::loggit(
  log_lvl = "DEBUG", "This message will be logged because the log level is ignored.",
  ignore_log_level = TRUE
)
loggit2::warning(
  "This warning message will not be logged, since .loggit = FALSE.",
  .loggit = FALSE
)

loggit2::set_log_level("ERROR")
loggit2::warning("This warning will not be logged, since the log level is set to ERROR.")
loggit2::message("This message will be logged, since .loggit = TRUE.", .loggit = TRUE)
loggit2::stop("This error message will be logged because the log level is set to ERROR.")

loggit2::with_loggit(log_level = "DEBUG", {
  base::message("This message will be logged because the log level is set to DEBUG.")
})

loggit2::set_log_level(old_log_lvl)
```

### Echo
Whether log messages should also be output to `stdout` can be controlled via the environment variable `ECHO_LOGGIT2` or the function `set_echo()`. Additionally, all logging functions and wrappers of the base R condition handler allow direct control via the `echo` argument.
```{r}
old_echo <- loggit2::set_echo(FALSE)
loggit2::message("This message will not be logged, but it will be output to the console.")
loggit2::message("This message will be logged and output to the console.", echo = TRUE)

loggit2::set_echo(TRUE, confirm = FALSE)
loggit2::message("This message will be logged and output to the console.")
loggit2::message("This message will be logged, but it will not be echoed.", echo = FALSE)

loggit2::with_loggit(echo = FALSE, {
  base::message("This message will be logged, but it will not be output to the console.")
})

loggit2::set_echo(old_echo)
```
### Timestamp format
The format of the timestamp can be controlled via the environment variable `TIMESTAMP_LOGGIT2` or the function `set_timestamp_format()`.

```{r}
old_ts <- loggit2::set_timestamp_format("%H:%M:%S")
loggit2::message("This message will be logged with a timestamp in the format HH:MM:SS.")

loggit2::set_timestamp_format(old_ts)
```

## Temporary Configuration
Depending on the use case, it may be useful to change temporary configurations, e.g., to log to a different file in a specific code block or to log a particular part of the code with more details (higher log level).

One way is to manually set and reset the configurations (e.g., by using `on.exit()`), but this is cumbersome and error-prone.

An alternative is to use `with_loggit()`. As mentioned above, almost all configurations can be adjusted directly in `with_loggit()`.

[^1]: Note: "Session-wide" means that (unless locally configured otherwise) calls to `loggit2` functions in other packages will use the same settings. Consider this when using `loggit2` in a package that will be used elsewhere.

```{r, include = FALSE}
options(old)
```
