---
title: "Why the Fork?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Why the Fork?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set( #nolint
  collapse = TRUE,
  comment = "#>"
)
```

Ideally, I would have preferred to further develop `{loggit}`, as I consider the concept highly commendable and worthy of
recognition. Unfortunately, this was not possible because the maintainer was unresponsive to messages or pull requests
at the time of the fork (the last release was over three years ago).
Additionally, CRAN's rules do not permit the takeover of the package without the maintainer's consent as
long as the package passes the checks.

Therefore, I ultimately decided to create a fork (`{loggit2}`) to address some fundamental issues and implement necessary
improvements. In the following, I will discuss the most pressing points.

## Call of conditions
In `{loggit}`, all "condition log handlers" (`loggit::message()`, `loggit::warning()`, `loggit::stop()`) presented themselves
as the call of the condition. This, combined with the fact that `{loggit}` does not allow passing a condition to the
"condition log handlers" (e.g., in `tryCatch()`), made it significantly more difficult to trace the origin of a condition.

``` r
loggit::message("This is a message")
#> {"timestamp": "2022-04-12T10:55:02-0500", "log_lvl": "INFO", "log_msg": "This is a message"}
#> This is a message
loggit::warning("This is a warning")
#> {"timestamp": "2022-04-12T10:55:02-0500", "log_lvl": "WARN", "log_msg": "This is a warning"}
#> Warning in loggit::warning("This is a warning"): This is a warning
loggit::stop("This is an error")
#> {"timestamp": "2020-05-31T20:59:33-0500", "log_lvl": "ERROR", "log_msg": "This is an error"}
#> Error in loggit::stop("This is an error"): This is an error
```
In `{loggit2}`, each condition handler shows the same call as its base equivalent.
```{r, error = TRUE}
base::message("This is another message")
loggit2::message("This is another message")

base::warning("This is another warning")
loggit2::warning("This is another warning")

base::stop("This is another error")
loggit2::stop("This is another error")
```

For further information and comparisons look [here](#further-comparisons).

## Missing `stopifnot()`
In `{loggit}`, there are no "condition log handlers" (i.e., functions that work like their base equivalents but also
write to the log) for `base::stopifnot()`. This makes the migration of existing code (without logs) unnecessarily complicated.

In `loggit2`, this has been added:
```{r, error = TRUE}
base::stopifnot("TRUE is not true" = TRUE, "This is an error" = 3L < 1L, "This is another error" = FALSE)
loggit2::stopifnot("TRUE is not true" = TRUE, "This is an error" = 3L < 1L, "This is another error" = FALSE)
```

## Log entries containing `:`

In `{loggit}`, log entries containing `: ` are not correctly parsed by `loggit::read_logs()`. In the best case, the
log messages are truncated after the `: `. See also [here](https://github.com/ryapric/loggit/issues/26).
This bug is fixed in `loggit2`.

## Clearing the log

In `{loggit}`, clearing the log with `loggit::rotate_logs(0L)` results in the log being unreadable by `loggit::read_logs()`,
even with new entries. See also [here](https://github.com/ryapric/loggit/issues/30). This bug is fixed in `loggit2`.

## Conclusion

In addition to these issues, there are several others, but these are the most pressing and can all be easily fixed.

Apart from the aforementioned bugs and missing features, there is significant untapped potential in the control settings (e.g., global settings regarding log levels) that can be implemented without complicating `{loggit}`.

All of these aspects will be addressed in `{loggit2}`. I welcome any suggestions for new features, bug reports, or even a direct pull request on either topic.

## Further comparisons

In `{loggit}`, each condition handler shows itself as the call of the condition.
``` r
f <- function() {
  loggit::message("This is another message")
  loggit::warning("This is another warning")
  loggit::stop("This is another error")
}

f()
#> {"timestamp": "2022-04-12T10:55:02-0500", "log_lvl": "INFO", "log_msg": "This is a message"}
#> This is a message
#> {"timestamp": "2022-04-12T10:55:02-0500", "log_lvl": "WARN", "log_msg": "This is a warning"}
#> Warning in loggit::warning("This is a warning"): This is a warning
#> {"timestamp": "2020-05-31T20:59:33-0500", "log_lvl": "ERROR", "log_msg": "This is an error"}
#> Error in loggit::stop("This is an error"): This is an error
```

In `{loggit2}`, each condition handler shows the same call as its `{base}` equivalent.
```{r, error = TRUE}
f <- function() {
  base::message("This is another message")
  base::warning("This is another warning")
  base::stop("This is another error")
}

f()

f <- function() {
  loggit2::message("This is another message")
  loggit2::warning("This is another warning")
  loggit2::stop("This is another error")
}

f()
```
